---
name: metrics_test_analysis
owner: vmandela99@gmail.com
metrics:
  avg_trp:
    title: payment consistency
    description: How well clients pay their loans
  nb_product:
    title: Type of product
    description: The type of product being sold
  nb_total_amount:
    title: Total amount paid
    description: Not adjusted for principal
  nb_cycle:
    title: Number of products
    description: Product series an individual has purchased
  nb_installments:
    title: Installment
    description: Number of installment periods
  avg_loan_principal:
    title: Mean principal amount (usd)
    description: The principal amount paid.
  avg_down_payment:
    title: Avg_downpayment
    description: Dinamic depending on the place
dimensions:
  product:
    title: The product name
    description: type of products being sold
  branch:
    title: County Name
    description: Operational counties that this product is sold
  status:
    title: Status of account
    description: Whether active or otherwise
---

## Description of the data

The table shows the sales made between Jan 2019 and March 2020

Each line is a sale (Account) with a product (e.g. LPG3) and a loan associated with the product (Total amount = Downpayment + Loan Principal)

Cycle indicates how many products the client has bought, 1 means this sale is the first one, 3 means is the third purchase for same client

TRP is a measure of how well the client is paying the loan, see below for definition


Each client belongs to a Group of at least five members

Each loan is a payment plan with a certain number of Instalments which are paid monthly

Each loan has a start date (Disbursed Date) and date of when it was fully paid (Final Payment Date), hence theoretical end date is Disbursed Date + N monthly Instalments


```{r Packages, include=FALSE}
library(tidyverse)                        #for manipulation
library(readxl)                           #for reading xlsx data
library(janitor)                          #for cleaning data

theme_set(theme_light())                  #friendly theme (easier to read)
```

# Explore

```{r pre-processing, include=FALSE}
sales <- readxl::read_xlsx("Book1.xlsx")    #get the data

filtered_sales <- sales %>% 
  janitor::clean_names() %>% 
  mutate_if(is.character, factor) %>% 
  mutate_at(vars(account), factor) %>% 
  select(-client_name) %>% 
    filter(!is.na(account))
  
skimr::skim(filtered_sales)
```

# impute the data

```{r}
library(naniar)

#look the amount of missing-ness
filtered_sales %>% 
  select(-c(client_id, dob, id_card, account, disbursed_date, group_id, final_payment_date)) %>% 
  gg_miss_upset()
```

```{r}
#library(recipes)

# imputed_df <- recipe( ~ cycle + trp, data = filtered_sales) %>% 
#   step_knnimpute(all_predictors()) %>% 
#   prep() %>% 
#   juice() %>% 
#   mutate(cycle = round(cycle))
# imputed_df %>% 
#   round(cycle)

```

# variable importance

Can you infer from this data in which order clients that bought more than one product make their purchases? Do they first buy an LPG and then a stove or a solar lamp?

# model

```{r}
library(lubridate)

sales_df <- 
  filtered_sales %>%
  transmute(
    more_than_one = case_when(cycle >= 2 ~ "Yes",
                              TRUE ~ "No"),
         age = year(Sys.Date()) - year(dob),
    #product = fct_lump(product, 8),
    gender, account, branch, product, loan_principal,
    status, trp) %>%  
  mutate_if(is.character, factor) %>% 
  na.omit() %>% 
  mutate_at(vars(account), as.integer)

skimr::skim(sales_df)
```


```{r}
library(tidymodels)

set.seed(123)

#workflow
sales_split <- initial_split(sales_df, strata = more_than_one)

sales_train <- training(sales_split)
sales_test <- testing(sales_split)

sales_rec <- recipe(more_than_one ~ .,
                    data = sales_df) %>% 
  update_role(account, new_role = "id") %>% 
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

sales_rec

sales_prep <- prep(sales_rec)
juiced <- juice(sales_prep)

```

```{r}
rf_spec <- rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger")


sales_wf <- workflow() %>%
  add_recipe(sales_rec) %>%
  add_model(rf_spec)

sales_wf
```

```{r}
library(vip)
library(ranger)

varible_importance <- rf_spec %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(more_than_one ~ .,
    data = juice(sales_prep) %>% select(-account))%>% 
      vip(geom = "point")


```



# Correlation


any correlation/pattern between age of client and choice of product bought?
repayment behaviour measured with TRP and choice of product bought?

any correlation/pattern between gender and choice of product bought?

```{r}

library(corrr)

cor_matrix <- sales_df %>% 
  #select(:has_bank_account) %>% 
  mutate_at(vars(more_than_one, gender, branch, product, status), as.integer) %>% 
  correlate() %>%
  rearrange() %>%
  shave() 

cor_matrix %>% 
  knitr::kable()

cor_matrix %>%
  rplot(shape = 15, colours = c("darkorange", "white", "darkcyan")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#Trend in payment of loans

how long do clients really take to fully pay their loans, what is the distribution? Any difference between products?


```{r}
filtered_sales %>% 
  filter(!is.na(final_payment_date)) %>% 
  mutate(days_to_fully_pay = day(final_payment_date) - day(disbursed_date)) %>% ggplot(aes(days_to_fully_pay, fill = product)) + 
  geom_density(alpha = 0.5) 
  
```

```{r}

library(tidymetrics)      #load the package

sales_summarized <- filtered_sales %>% 
  filter(product %in% c("LPG3", "LPG2", "CNVS", "P400",
                     "JIKO", "BORA",
                     "SKH"),
         !is.na(final_payment_date)) %>%
  rename(date = final_payment_date) %>%
  mutate(date = as.Date(date)) %>% 
  cross_by_dimensions(product, branch, status) %>%
  cross_by_periods(c("day", "week", "month", "year"),
                   windows = 28) %>%
  summarize(avg_trp = mean(trp, na.rm = T), 
            nb_product = n_distinct(product),
            nb_total_amount = sum(total_amount),
            nb_cycle = sum(cycle),
            nb_installments = sum(installments),
            avg_loan_principal = mean(loan_principal),
            avg_down_payment = mean(down_payment)
            ) %>%    #theatre
  ungroup()

```

```{r}
use_metrics_scaffold(sales_summarized)       #get the metrics details
sales_metrics <- create_metrics(sales_summarized)
sales_metrics

saveRDS(sales_metrics, "sales_metrics.rds")
```

```{r}
library(shinymetrics)

preview_metric(sales_metrics$test_analysis_avg_trp)
preview_metric(sales_metrics$test_analysis_nb_product)
preview_metric(sales_metrics$test_analysis_nb_total_amount)
preview_metric(sales_metrics$test_analysis_nb_cycle)
preview_metric(sales_metrics$test_analysis_nb_installments)
preview_metric(sales_metrics$test_analysis_avg_loan_principal)
preview_metric(sales_metrics$test_analysis_avg_down_payment)

```






















